\documentclass[../main.tex]{subfiles}
\graphicspath{{\subfix{../images/}}}

\begin{document}

\newpage
\chapter{BACKGROUND}

\section{FDTD}

\subsection{Yee's algorithm}

The finite-difference time-domain (FDTD) method\cite{ufdtd} is used to aproximate electromagnetic fields over time.
It consists on a pair of ``update equations'' based on discretized versions of Ampère's law and Faraday's law that are called $t$ times.
Each equation takes the previous iteration's fields and some constants from the material properties as inputs and sets the current iteration electric or magnetic field.

The core of Fdtd Lucuma's simulation loop at each time step consists of the previously mentioned two equations, one or more additive electric sources and the boundary conditions of each field and component.
Fdtd Lucuma currently only has the gaussian source as is sole source implementation and the Absorbing boundary condition as the boundary implementation.

\begin{figure}[H]
	\centering
	\begin{tikzpicture}[
			transform shape,
			vertex/.style={draw, ellipse, thick},
		]
		\node[vertex] (A) at (0,0) {Update magnetic $\{x,y,z\}$ fields};
		\node[vertex] (B) [below=of A] {Update electric $\{x,y,z\}$ fields};
		\node[vertex] (C) [below=of B] {Apply sources};
		\node[vertex] (D) [below=of C] {Boundary conditions};

		\path[-{Triangle[length=5]}, thick]
			(A) edge (B)
			(B) edge (C)
			(C) edge (D)
			(D.west) edge[bend left] (A.west)
		;

	\end{tikzpicture}
	\caption{Simulation step sequence}
	\label{fig:fdtd_lucuma_step}
\end{figure}

\subsection{3D}

\subsubsection{Variables}
\begin{itemize}
	\item $x,y,z,t$: 3D space and time.
	\item $q \in \mathbb{R}$: Time index.
	\item $[m,n,p] \in \mathbb{R}^3:$ Spatial index.
	\item $H_i^q$: Magnetic field component at time $q$. $i \in \{x,y,z\}$
	\item $E_i^q$: Electric field component at time $q$. $i \in \{x,y,z\}$
	\item $\sigma$: Electric conductivity of a material.
	\item $\sigma_m$: Magnetic conductivity of a material.
	\item $\mu$: Permeability of a material.
	\item $\epsilon$: Permittivity of a material.
	\item $C\{e,h\}\{x,y,z\}\{e,h\}$: Coefficients.
	\item $Cr$: Courant number.
\end{itemize}

\directlua{
function fmt(x)
	if x == "" then
		return ""
	end

	local n = tonumber(x)

	if n == 0.5 then
		return "+\\frac{1}{2}"
	elseif n == -0.5 then
		return "-\\frac{1}{2}"
	elseif n > 0 then
		if n == math.floor(n) then
			return "+" .. tostring(math.floor(n))
		else
			return "+" .. tostring(n)
		end
	elseif n < 0 then
		if n == math.floor(n) then
			return tostring(math.floor(n))
		else
			return tostring(n)
		end
	else
		return ""
	end
end

function flatfmt(n, label)
	if n == "" then
		return label
	end

	return n
end

function coords(a,b,c)
	tex.sprint(
		"\\left[m" .. fmt(a) .. ",n" .. fmt(b) .. ",p" .. fmt(c) .. "\\right]"
	)
end

function flatcoords(a,b,c)
	return "\\left[" .. flatfmt(a, "m") .. "," .. flatfmt(b, "n") .. "," .. flatfmt(c, "p") .. "\\right]"
end

function abcComponent(c, coords, deltaCoords)
	tex.sprint("\\ABCComponent{"..c.."}{"..coords.."}{"..deltaCoords.."}\\\\")
end

function side2CoordsCommon(c, i)
	if c == "x" then
		return flatcoords(i,"","")
	elseif c == "y" then
		return flatcoords("",i,"")
	elseif c == "z" then
		return flatcoords("","",i)
	end
end

function side2Coords(c, side)
	local i = (function() if side == "top" then return c else return "1" end end)()

	return side2CoordsCommon(c, i)
end

function side2DeltaCoords(c, side)
	local i = (function() if side == "top" then return c .. "-1" else return "2" end end)()

	return side2CoordsCommon(c, i)
end

function other(c, c2)
	if c == c2 then
		return
	end

	abcComponent(c2, side2Coords(c, "bottom"), side2DeltaCoords(c, "bottom"))
	abcComponent(c2, side2Coords(c, "top"), side2DeltaCoords(c, "top"))
end

function otherComponents(c)
	other(c, "x")
	other(c, "y")
	other(c, "z")
end

}

\newcommand{\Sc}{%
	\ensuremath{\frac{Cr}{\sqrt{\mu*\epsilon}}}%
}

\newcommand{\ABCComponent}[3]{%
	\ensuremath{E_{#1}^{q+1}#2 &= E_{#1}^q#2 + \frac{\Sc-1}{\Sc+1}*\left(E_{#1}^{q}#3-E_{#1}^{q}#2\right)}%
}

\newcommand{\coords}[3]{%
	\ensuremath{\directlua{
		coords(
			"\luaescapestring{#1}",
			"\luaescapestring{#2}",
			"\luaescapestring{#3}"
		)
	}}%
}

\newcommand{\flatcoords}[3]{%
	\ensuremath{\directlua{
		flatcoords(
			"\luaescapestring{#1}",
			"\luaescapestring{#2}",
			"\luaescapestring{#3}"
		)
	}}%
}
\newcommand{\otherComponents}[1]{%
	\ensuremath{\directlua{
		otherComponents(
			"\luaescapestring{#1}"
		)
	}}%
}

\newcommand{\HHcoeff}[1]{%
	\ensuremath{\underbrace{\frac{1-\frac{\sigma_m\Delta_t}{2\mu}}{1+\frac{\sigma_m\Delta_t}{2\mu}}}_{Ch#1h}}%
}

\newcommand{\HEcoeff}[1]{%
	\ensuremath{\underbrace{\frac{1}{1+\frac{\sigma_m\Delta_t}{2\mu}}}_{Ch#1e}}%
}

\newcommand{\EEcoeff}[1]{%
	\ensuremath{\underbrace{\frac{1-\frac{\sigma\Delta_t}{2\epsilon}}{1+\frac{\sigma\Delta_t}{2\epsilon}}}_{Ce#1e}}%
}

\newcommand{\EHcoeff}[1]{%
	\ensuremath{\underbrace{\frac{1}{1+\frac{\sigma\Delta_t}{2\epsilon}}}_{Ce#1h}}%
}

\newcommand{\CommonDelta}[2]{%
	\ensuremath{\frac{\Delta_t}{#1\Delta_{#2}}}%
}

\newcommand{\EDelta}[1]{%
	\ensuremath{\CommonDelta{\mu}{#1}}%
}

\newcommand{\HDelta}[1]{%
	\ensuremath{\CommonDelta{\epsilon}{#1}}%
}

\newcommand{\Hupdate}[8]{%
	\ensuremath{H_{#1}^{q+\frac{1}{2}} #4 &=%
	\HHcoeff{#1} H_{#1}^{q-\frac{1}{2}}#4 \\%
	+& \HEcoeff{#1} \biggl(\EDelta{#3}\left\{E_{#2}^q#5-E_{#2}^q#6\right\} \\%
		-&\EDelta{#2}\left\{E_{#3}^q#7-E_{#3}^q#8\right\}%
	\biggl)}%
}

\newcommand{\Eupdate}[8]{%
	\ensuremath{E_{#1}^{q+1} #4 &=%
	\EEcoeff{#1} E_{#1}^{q}#4 \\%
	+& \EHcoeff{#1} \biggl(\HDelta{#3}\left\{H_{#2}^{q+\frac{1}{2}}#5-H_{#2}^{q+\frac{1}{2}}#6\right\} \\%
		-&\HDelta{#2}\left\{H_{#3}^{q+\frac{1}{2}}#7-H_{#3}^{q+\frac{1}{2}}#8\right\}%
	\biggl)}%
}

\subsubsection{Update magnetic field components}
\begin{equation}
\begin{split}
	\Hupdate{x}{y}{z}{\coords{}{0.5}{0.5}}{\coords{}{0.5}{1}}{\coords{}{0.5}{}}{\coords{}{1}{0.5}}{\coords{}{}{0.5}}
\end{split}
\end{equation}
\begin{equation}
\begin{split}
	\Hupdate{y}{z}{x}{\coords{0.5}{}{0.5}}{\coords{1}{}{0.5}}{\coords{}{}{0.5}}{\coords{0.5}{}{1}}{\coords{0.5}{}{}}
\end{split}
\end{equation}
\begin{equation}
\begin{split}
	\Hupdate{z}{x}{y}{\coords{0.5}{0.5}{}}{\coords{0.5}{1}{}}{\coords{0.5}{}{}}{\coords{1}{0.5}{}}{\coords{}{0.5}{}}
\end{split}
\end{equation}


\subsubsection{Update electric field components}
\begin{equation}
\begin{split}
	\Eupdate{x}{z}{y}{\coords{}{0.5}{}}{\coords{0.5}{0.5}{}}{\coords{0.5}{-0.5}{}}{\coords{0.5}{}{0.5}}{\coords{0.5}{}{-0.5}}
\end{split}
\end{equation}
\begin{equation}
\begin{split}
	\Eupdate{y}{x}{z}{\coords{}{0.5}{}}{\coords{}{0.5}{0.5}}{\coords{}{0.5}{-0.5}}{\coords{0.5}{0.5}{}}{\coords{-0.5}{0.5}{}}
\end{split}
\end{equation}
\begin{equation}
\begin{split}
	\Eupdate{z}{y}{x}{\coords{}{}{0.5}}{\coords{0.5}{}{0.5}}{\coords{-0.5}{}{0.5}}{\coords{}{0.5}{0.5}}{\coords{}{-0.5}{0.5}}
\end{split}
\end{equation}

\subsubsection{Gaussian source}
\begin{equation}
\begin{split}
	E_x^{q+1}\left[m,n,p\right] &= E_x^{q}\left[m,n,p\right] + \exp\left(-\left(\frac{q}{\sigma}\right)^2\right)
\end{split}
\end{equation}

\subsubsection{Absorbing boundary condition}

\newcommand{\EABC}[1]{%
\begin{equation}%
\begin{split}%
	\otherComponents{#1}%
\end{split}%
\end{equation}%
}

\EABC{x}
\EABC{y}
\EABC{z}


\section{Julia + CUDA}
Julia\cite{juliaspec} is a dynamically typed programming language mainly used for scientific computing.
Moreover, the \texttt{CUDA.jl}\cite{besard2018juliagpu} package allows it to interface with Nvidia GPUs.
Here is an example, a kernel that sums two arrays value-wise and sets the results in a third array:
\inputminted[bgcolor=codeBg, tabsize=4, breaklines=true, linenos=true]{julia}{\subfix{add.jl}}

\section{Vulkan}
Vulkan\cite{vulkanspec} is a low-level C API for graphics and compute in GPUs.
Unlike CUDA, which only runs on Nvidia hardware natively\footnote{With ZLUDA is posible to run CUDA programs in Radeon hardware.}, it has a wider range of compatible GPU hardware\cite{vulkandb}.

In OpenGL, Vulkan's predecesor, the shader source file can be passed directly to the driver, but Vulkan requires
for each shader to be compiled into a spirv file, loaded into a shader module\footnote{From Vulkan 1.4 onwards the \texttt{*.spv} file path can be passed directly into the creation of the pipeline.} and then finally used as part of a compute pipeline before it can be used.
For that purpose a shader compiler is needed.
Like Google's \texttt{glslc}\cite{shaderc} or Shader Slang's \texttt{slangc}\cite{10.1145/3197517.3201380} (which is the one that we used in this project).


\begin{figure}[H]
	\centering
	\begin{tikzpicture}[
			transform shape,
			vertex/.style={draw, ellipse, thick},
			scale=0.9
		]
		\node[vertex] (A) at (0,0) {\texttt{*.spv}};
		\node[vertex] (B) [above left=of A] {\texttt{*.slang}};
		\node[vertex] (C) [below left=of A] {\texttt{*.glsl}};
		\node[vertex] (F) [left=of A] {\texttt{*.hlsl}};
		\node[vertex] (D) [right=of A] {\texttt{VkShaderModule}};
		\node[vertex] (E) [right=of D] {\texttt{VkPipeline}};

		\path[-{Triangle[length=5]}, thick]
			(B) edge (A)
			(C) edge (A)
			(F) edge (A)
			(A) edge (D)
			(D) edge (E)
			(A) edge[bend left, dashed] node[label=above:Since Vulkan 1.4\cite{vulkan14}] {} (E)
		;

	\end{tikzpicture}
	\caption{Vulkan compute pipeline creation}
	\label{fig:vulkan_compute_pipeline}
\end{figure}

Using Vulkan Compute Shaders requires writing a considerable amount of boilerplate.
Therefore the complete version of the sum two arrays example is in the Github repository\footnote{\tiny{\url{https://github.com/fdtd-lucuma/fdtd-lucuma/blob/53fc40cde634b7360941ea330caa1e1fe68697a8/src/services/compute.cpp\#L34}}}:

\inputminted[bgcolor=codeBg, tabsize=4, breaklines=true, linenos=true]{hlsl}{\subfix{add.slang}}

\inputminted[bgcolor=codeBg, tabsize=4, breaklines=true, linenos=true]{cpp}{\subfix{add.cpp}}


\end{document}
